---
title: "Crime Rates By County"
author: "Moises Evangelista"
date: "January 19, 2016"
output: pdf_document
fontsize: 11pt
header-includes:
- \usepackage{palatino}
- \renewcommand{\familydefault}{\sfdefault} % sans serif
- \fontfamily{ppl}\selectfont
---

This report uses data from the FBI UCR tables to get known crimes rates by year for different cities in California. The goal is to determine crime rate changes for California counties before and after AB109 was implemented.sa

Data sources:

1.[City crimes rates from FIB table 8](http://www.fbi.gov/about-us/cjis/ucr/crime-in-the-u.s/)

2.[City of California shapefile](http://www.dot.ca.gov/hq/tsip/gis/datalibrary/Metadata/cities.html)

3.[Counties shape file from U.S. Census](https://www.census.gov/geo/maps-data/data/cbf/cbf_counties.html)

\vspace{5mm}

The FBI data has the following disclaimer:

Caution against ranking
Figures used in this Report were submitted voluntarily by law enforcement agencies throughout the country. Individuals using these tabulations are cautioned against drawing conclusions by making direct comparisons between cities. Comparisons lead to simplistic and/or incomplete analyses that often create misleading perceptions adversely affecting communities and their residents. Valid assessments are possible only with careful study and analysis of the range of unique conditions affecting each local law enforcement jurisdiction. It is important to remember that crime is a social problem and, therefore, a concern of the entire community. In addition, the efforts of law enforcement are limited to factors within its control. The data user is, therefore, cautioned against comparing statistical data of individual agencies. Further information on this topic can be obtained in Uniform Crime Reporting Statistics:  Their Proper Use.

In other words, comparing cities among other cities is not a good idea, but comparing a city within itself by year is OK.

\newpage

```{r echo=FALSE, message=FALSE, include=FALSE}

rm(list = ls(all = TRUE)) #start with empty workspace

startTime <- Sys.time()

# set global chunk options

knitr::opts_chunk$set(cache = TRUE, echo = FALSE, include = FALSE)

doInstall <- FALSE # Change to TRUE if you do want packages installed.
toInstall <- c("knitr", "rvest", "NbClust", "cluster", "foreign", "dplyr", 
               "ggplot2", "scales", "RCurl", "RODBC", "xtable",
               "benford.analysis", "strucchange", "forecast", "tidyr",
               "RColorBrewer")
if(doInstall){install.packages(toInstall, repos = "http://cran.us.r-project.org")}
lapply(toInstall, library, character.only = TRUE)

setwd("~/CrimeRatesByCounty/") # set the working directory
list.files() # see whats there

```

```{r downloadData, echo=FALSE, include=FALSE, eval=TRUE}

# Download FBI Data

theurlPart1 <- "http://www.fbi.gov/about-us/cjis/ucr/crime-in-the-u.s/"

theurlPart214 <- "2014/crime-in-the-u.s.-2014/tables/table-8/table-8-by-state/"
theurlPart314 <- "Table_8_Offenses_Known_to_Law_Enforcement_by_California_by_City_2014.xls"

theurlPart213 <- "2013/crime-in-the-u.s.-2013/tables/table-8/table-8-state-cuts/"
theurlPart313 <- "table_8_offenses_known_to_law_enforcement_california_by_city_2013.xls"

theurlPart212 <- "2012/crime-in-the-u.s.-2012/tables/8tabledatadecpdf/"
theUrlPart213 <- "table-8-state-cuts/table_8_offenses_known_to_law_enforcement_by_california_by_city_2012.xls"

theurlPart211 <- "2011/crime-in-the-u.s.-2011/tables/table8statecuts/"
theUrlPart311 <- "table_8_offenses_known_to_law_enforcement_california_by_city_2011.xls"

theurlPart210 <- "2010/crime-in-the-u.s.-2010/tables/table-8/10tbl08ca.xls"

theurl2009 <- "http://www2.fbi.gov/ucr/cius2009/data/table_08_ca.html"

# set col names

table8ColNamesCurrent <- c("City","Population","Violent crime",
                           "Murder and nonnegligent manslaughter","Rape (revised definition)1",
                           "Rape (legacy definition)2","Robbery","Aggravated assault",
                           "Property crime","Burglary","Larceny- theft",
                           "Motor vehicle theft","Arson")

table8ColNames <- c("City","Population","Violent crime",
                           "Murder and nonnegligent manslaughter",
                           "Rape (legacy definition)2","Robbery",
                           "Aggravated assault","Property crime",
                           "Burglary","Larceny- theft","Motor vehicle theft",
                           "Arson")

# import the data for 2013
# tables <- readHTMLTable(paste0(theurlPart1,theurlPart213,theurlPart313))
# n.rows <- unlist(lapply(tables, function(t) dim(t)[1]))  # get the table with the highest number of rows

content2014 <- read_html(paste0(theurlPart1,theurlPart214,theurlPart314)) %>%
    html_table(fill = TRUE) %>%
    .[[2]] # table 3 only

colnames(content2014) <- table8ColNamesCurrent

content2014$Type <- "CA2014"

str(content2014)
summary(content2014)

content2013 <- read_html(paste0(theurlPart1,theurlPart213,theurlPart313)) %>%
    html_table(fill = TRUE) %>%
    .[[3]] # table 3 only

colnames(content2013) <- table8ColNamesCurrent

str(content2013)
summary(content2013)

# add a variable to ID records

content2013$Type <- "CA2013"

str(content2013)
summary(content2013)

# import the data for 2012

content2012 <- read_html(paste0(theurlPart1, theurlPart212, theUrlPart213)) %>%
    html_table(fill = TRUE) %>%
    .[[3]] # table 3 only

# add column names

colnames(content2012) <- table8ColNames

# add a variable to ID records

content2012$Type <- "CA2012"

str(content2012); summary(content2012)

content2011 <- read_html(paste0(theurlPart1, theurlPart211, theUrlPart311)) %>%
    html_table(fill = TRUE) %>%
    .[[3]] # table 3 only

# add column names

colnames(content2011) <- table8ColNames 

# add a variable to ID records

content2011$Type <- "CA2011"

str(content2011); summary(content2011)

content2010 <-  read_html(paste0(theurlPart1, theurlPart210)) %>%
    html_table(fill = TRUE) %>%
    .[[3]] # table 3 only

# add column names

colnames(content2010) <- table8ColNames 

# add a variable to ID records

content2010$Type <- "CA2010"

str(content2010); summary(content2010)

content2009 <- read_html("http://www2.fbi.gov/ucr/cius2009/data/table_08_ca.html") %>%
    html_table(fill = TRUE) %>%
    .[[1]] # table 3 only

# add column names

colnames(content2009) <- table8ColNames

# add a variable to ID records

content2009$Type <- "CA2009"

str(content2009); summary(content2009)

# combine all the records

comboRecs <- rbind(content2013, content2014) %>%
  select(-`Rape (revised definition)1`) %>%
  rbind(content2009, content2010, content2011, content2012)

# make numeric colms

str(comboRecs)

comboRecs <- comboRecs %>%
  mutate(Population = as.numeric(gsub(",","", 
                                      Population)),
         `Violent crime` = as.numeric(gsub(",","", 
                                           `Violent crime`)),
         `Murder and nonnegligent manslaughter` = as.numeric(gsub(",","", 
                                                                  `Murder and nonnegligent manslaughter`)),
         `Rape (legacy definition)2` = as.numeric(gsub(",","", 
                                                       `Rape (legacy definition)2`)),
         Robbery = as.numeric(gsub(",","", 
                                   Robbery)),
         `Aggravated assault` = as.numeric(gsub(",","", 
                                                `Aggravated assault`)),
         `Property crime` = as.numeric(gsub(",","", 
                                            `Property crime`)),
         `Burglary` = as.numeric(gsub(",","", 
                                      `Burglary`)),
         `Larceny- theft` = as.numeric(gsub(",","", 
                                            `Larceny- theft`)),
         `Motor vehicle theft` = as.numeric(gsub(",","",
                                                 `Motor vehicle theft`)),
         Arson = as.numeric(gsub(",","", Arson)))

str(comboRecs)

# get the rates for each type of crime and violent and property crimes

comboRecs <- comboRecs %>%
  mutate(ViolentCrimeRate = (`Violent crime`/Population)*100000,
         MurderAndNonnegligentManslaughterRate = (`Murder and nonnegligent manslaughter`/Population)*100000,
         RapeLegacyDefinitionRate = (`Rape (legacy definition)2`/Population)*100000,
         RobberyRate = (Robbery/Population)*100000,
         AggravatedAssaultRate = (`Aggravated assault`/Population)*100000,
         PropertyCrimeRate = (`Property crime`/Population)*100000,
         BurglaryRate = (Burglary/Population)*100000,
         LarcenyTheftRate = (`Larceny- theft`/Population)*100000,
         MotorVehicleTheftRate = (`Motor vehicle theft`/Population)*100000,
         ArsonRate = (Arson/Population)*100000)

# some city names have numbers

comboRecs$City <-  gsub("[[:digit:]]", "", comboRecs$City)

# change the values of some cities 

comboRecs$City <-  gsub("Carmel", "Carmel-by-the-Sea", comboRecs$City)
comboRecs$City <-  gsub("Rancho Santa Margarit", "Rancho Santa Margarita", comboRecs$City)
comboRecs$City <-  gsub("Rancho Santa Margaritaa", "Rancho Santa Margarita", comboRecs$City)

```



```{r downloadShapeFiles, eval=TRUE, include=FALSE}

# city shape file 
# http://www.dot.ca.gov/hq/tsip/gis/datalibrary/Metadata/cities.html

library(maptools)

# create an interactive map like this

# http://spatial.ly/2013/12/introduction-spatial-data-ggplot2/

# US Counties

download.file('http://www2.census.gov/geo/tiger/GENZ2014/shp/cb_2014_us_county_500k.zip', 'county.zip',
              quiet = TRUE)

temp <- unzip('county.zip')

Counties <- readShapePoly("cb_2014_us_county_500k")

Counties@data$id <- rownames(Counties@data)

head(Counties@data)

# filter counties in Calif

CaCounties <-  subset(Counties, 
                      STATEFP == "06")

head(CaCounties@data)

tail(CaCounties@data)

# CAlifornia Cities

temp <- tempfile()


download.file("http://www.dot.ca.gov/hq/tsip/gis/datalibrary/zip/Boundaries/Cities2015.zip",
               quiet = TRUE, 
              destfile = temp)

con <- unzip(temp)

# take a look at the files on the data
con

califCities <- readShapePoly("./Cities2015")

head(califCities@data)

head(califCities@polygons)

frtfcalifCities <- fortify(califCities, region = "NAME")

frtfCaCounties <- fortify(CaCounties, region = "NAME")

###
### http://stackoverflow.com/questions/19791210/r-ggplot2-merge-with-shapefile-and-csv-data-to-fill-polygons

ggplot(frtfcalifCities, aes(long, lat, group = group)) +
  geom_polygon(aes(),  fill = "grey40", 
               colour = "grey90", alpha = .7, size = .05) +
  coord_map() +
  theme_minimal() 

ggplot(frtfCaCounties, aes(long, lat, group = group)) +
  geom_polygon(aes(),  fill = "blue", colour = "red", 
               alpha = .2, size = .05) +
  coord_map() +
  theme_minimal() 

ggplot(frtfcalifCities, aes(long, lat, group = group)) +
  geom_polygon(aes(),  fill = "grey40", 
               colour = "blue", alpha =.7, size = .05) +
  geom_path(data=frtfCaCounties, aes(long,lat, group = group,
                                     fill=hole), color = "red",
            alpha = .2, size = 0.3) +
  coord_map() +
  theme_minimal() 

##  clip the cities that are not withing the county layer

library(rgeos)

head(califCities@data)

califCities@data$id <- rownames(califCities@data)

citiesInterset <- gIntersection(califCities, CaCounties, 
                     byid = TRUE,
                     drop_lower_td = TRUE)

frtfCitiesInterset <- fortify(citiesInterset, region = "NAME")

# get the id field

head(califCities@data)
head(frtfCitiesInterset)

frtfCitiesInterset <- frtfCitiesInterset %>% 
  separate(id, into = c("NewID1", "NewID2"), sep = " ")

frtfCitiesInterset <- left_join(frtfCitiesInterset, 
                                califCities@data, 
                                by = c('NewID1' = 'id'))

ggplot(frtfCitiesInterset, aes(long, 
                               lat, 
                               group=group)) +
  geom_polygon(aes(),
               fill = "grey40",
               colour = "blue",
               alpha = .7,
               size = .05) +
  geom_polygon(data = frtfCaCounties, aes(long,
                                        lat,
                                        group = group), 
            fill = "khaki",
            color = "red",
            alpha = .2,
            size = 0.1) +
  coord_map() +
  theme_minimal() 

# merget the pas with the set of table8 crimes

head(frtfCitiesInterset)

comboCities <- left_join(frtfCitiesInterset, 
                                comboRecs, 
                                by = c('NAME' = 'City')) %>%
  filter(!is.na(Type)) %>%
  unique()

# take a look at those cities not on the map layer

anti_join(frtfCitiesInterset,
          comboRecs,
          by = c('NAME' = 'City')) %>%
  select(County, NAME) %>%
  distinct()

anti_join(comboRecs,
          frtfCitiesInterset,
          by = c('City' = 'NAME')) %>%
  select(City) %>%
  distinct()

# get the limits for each county

longLimits <- frtfCaCounties %>%
  group_by(id) %>%
  select(long, id) %>%
  unique() %>%
  arrange(long) %>%  ## optional
  filter(long %in% range(long)) %>%
   mutate(Number = 1,
          cumsum = cumsum(Number)) %>%
  select(long, id, cumsum) %>%
  mutate(cumsum = ifelse(cumsum == 1, 
                         "longMax", "LongMin")) %>%
  spread(cumsum, long)

latLimits <- frtfCaCounties %>%
  group_by(id) %>%
  select(lat, id) %>%
  unique() %>%
  arrange(lat) %>%  ## optional
  filter(lat %in% range(lat)) %>%
   mutate(Number = 1,
          cumsum = cumsum(Number)) %>%
  select(lat, id, cumsum) %>%
  mutate(cumsum = ifelse(cumsum == 1, 
                         "latMax", "LatMin")) %>%
  spread(cumsum, lat)


# centroids.df <- as.data.frame(coordinates(comboCities))
# names(centroids.df) <- c("Longitude", "Latitude") 

comboCities %>%
  filter(NAME != 'Vernon' &
           NAME !='Industry') %>%
  # select(NAME) %>%
  # distinct() %>% arrange((NAME))
  ggplot( aes(long, lat, group = group, fill = ViolentCrimeRate)) +
  geom_polygon(aes(),
               # fill="grey40",
               colour = "grey90",
               alpha = .7,
               size = .05) +
  # geom_text(aes(label = NAME, x = Longitude, y = Latitude), size = 1) +
  scale_fill_gradientn(name = "Rate",
                       colours = rainbow(7),
                       limits = c(0,2000)) +
  geom_polygon(data = frtfCaCounties, aes(long,
                                          lat,
                                          group = group), 
               fill = "khaki",
               color = "red",
               alpha = .2,
               size = 0.1) +
  coord_map() +
  theme_minimal() + 
  facet_wrap(~ Type)

```

```{r createbenfordAnalysis, eval=TRUE, include=FALSE}

benfordAnalysis <- comboCities %>%
  select(NAME, County, Year = Type,
         `Murder and nonnegligent manslaughter`:`Aggravated assault`,
         Burglary:Arson) %>%
  unique() %>%
  gather(Metric, Count, -c(NAME, County, Year)) %>%
  filter(Year == "CA2014")
```
```{r createbenfordAnalysis1, eval=TRUE, include=FALSE}
cp <- benford(benfordAnalysis$Count, 2, sign = "both")
# cp #prints
# plot(cp) #plots

```
```{r createbenfordAnalysis2, eval=TRUE, include=FALSE}

head(suspectsTable(cp), 10) #prints the digits by decreasing order of discrepancies
#gets observations of the 2 most suspicious groups
suspects <- getSuspects(cp, benfordAnalysis, how.many = 2) %>%
  na.omit()

nrow(suspects)/nrow(benfordAnalysis)
```
```{r createbenfordAnalysis3, eval=FALSE, include=FALSE}

benfordAnalysisPop <- comboCities %>%
  select(NAME, County, Year = Type,
         Pop2010) %>%
  # mutate(Pop2010 = as.numeric(Pop2010)) %>%
  unique() %>%
  gather(Metric, Count, -c(NAME, County, Year)) %>%
  filter(Year == "CA2014")

cp <- benford(benfordAnalysisPop$Count, 2, sign = "both")
cp #prints
plot(cp) #plots

head(suspectsTable(cp), 10) #prints the digits by decreasing order of discrepancies
#gets observations of the 2 most suspicious groups
suspectsPop <- getSuspects(cp, benfordAnalysisPop, how.many = 2) %>%
  na.omit()

# percent of pop counts not deviating from bedfords law

nrow(suspectsPop)/nrow(benfordAnalysis)

```

```{r exportTheData, eval=FALSE}

# export the data

dir.create("./data")

comboCities %>%
  select(long, lat, NAME, County, Type, group, 
  ends_with("Rate")) %>%
  write.table(file = "./data/comboCities.txt", 
  sep = "\t", na = "0", row.names = FALSE)

frtfCaCounties %>%
  select(long, lat, id, group) %>%
  write.table(file = "./data/frtfCaCounties.txt", 
  sep = "\t", na = "0", row.names = FALSE)

left_join(longLimits, latLimits) %>%
  write.table( file = "./data/countyLimits.txt", 
               sep = "\t", na = "0", 
               row.names = FALSE)

suspects %>%   write.table(file = "./data/suspects.txt",
sep = "\t", na = "0", row.names = FALSE)

```

```{r plotCityRatesbyYear, fig.height=10,  eval=TRUE, include=TRUE}

colourCount = length(unique(comboCities$Type))
getPalette = colorRampPalette(brewer.pal(8, "Accent"))

comboCities %>%
  select(ViolentCrimeRate, NAME, Type, County) %>% 
  filter(County == "Los Angeles") %>%
  distinct() %>% 
  ggplot(aes(ViolentCrimeRate, NAME)) +
  geom_point(aes(colour = factor(Type),
                 shape = factor(Type))) +
  scale_x_continuous(trans = log_trans(),
                     breaks = c(100,1000,10000,
                                50000),
                     labels = comma_format()) +
  scale_shape_manual(name = "Rate \nper year",
                     values = c(0:5)) +
  scale_color_manual(values = getPalette(colourCount),
                     name = "Rate \nper year",
                     guide = guide_legend(keywidth = .5,
                                          keyheight = .5,
                                          direction = "vertical",
                                          title.position = "left",
                                          label.position = "right", 
                                          label.hjust = 0, 
                                          label.vjust = 0, 
                                          ncol = 3, 
                                          byrow = TRUE,
                                          title.theme = element_text(size = 5, 
                                                                     colour = "black", 
                                                                     angle = 00),
                                          label.theme = element_text(size = 6, 
                                                                     colour = "black", 
                                                                     angle = 00))) +
  theme_minimal() +
  xlab("Crime Rate per 100,000 residents (log scale)") +
  ylab("City") +
  theme(axis.text.x  = element_text(angle = 00,
                                    vjust = 0.5, size = 8), 
        axis.text.y  = element_text(angle = 0,
                                    vjust = 0.5, size = 8),
        legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

```

```{r echo=FALSE}

# TODO
# create bendford analysis -- done
# add city labels to the map
# add zoom feature to map
# add segment that compares crime rate changes since 2011
# add good limits for map fill colors

```


